### 2022-07-15 draft.tex
Deploying to gh-pages from @ graciellehigino/ms_range_interactions@2fa2e76f994bf8a152eb7a024c46fe00746dd3e5 🚀

Finding a species in a certain location is like finding an encrypted
message that traveled through time. It carries the species' evolutionary
history, migration patterns, as well as any direct and indirect effects
generated by other species (some of which we may not even know exist).
Ecologists have been trying to decode this message with progressively
more powerful tools, from their field notes to highly complex
computational algorithms. However, to succeed in this challenge it is
important to have the right clues in hand. There are many ways we can be
misled by data - or the lack of it: taxonomic errors (e.g., due to
updates in the taxonomy of a species), geographic inaccuracy (e.g.,
approximate coordinates or lack of documentation about their accuracy),
or sampling biases (e.g.~data clustered near roads or research centers)
(Ladle and Hortal 2013; Hortal et al. 2015; Poisot et al. 2021). One way
to identify - and potentially fix - these errors is to combine many
different pieces of information about the occurrence of a species, so
agreements and mismatches can emerge. Although previous studies have
combined different types of occurrence data to measure the accuracy of
datasets (Hurlbert and Jetz 2007; Hurlbert and White 2005;
\textbf{Ficetola2014EvaRob?}), none have used different types of
information so far (i.e., ecological characteristics other than
geographical distribution). Here we suggest jointly analysing species
occurrence (range maps and point occurrences) and ecological
interactions to identify mismatches between datasets.

@@

It has been demonstrated that the agreement between range maps and point
data varies geographically (Hurlbert and Jetz 2007; Hurlbert and White
2005; \textbf{Ficetola2014EvaRob?}). Adding ecological interaction data
to this comparison might help to elucidate where these (dis)agreements
are more likely to be true and which dataset better represent the actual
distribution of species. In this context, we elaborate a method that
allows us to refine distribution data (more precisely range maps) based
on interaction data, considering the basic assumption that predators can
only be present in regions where they are connected to at least one
herbivore - and thus indirectly connected to primary producers. We used
a Serengeti food web dataset (Baskerville et al. 2011) (which comprises
carnivores, herbivores, and plants from Tanzania) to demonstrate how a
mismatch between occurrence and interaction data can highlight
significant uncertainty areas in IUCN range maps. Finally, we add the
GBIF occurrence points for the Serengeti species to the investigation,
discuss the mechanisms that can lead to the lack of agreement between
data, and build from that a vision for the next steps, reinforcing the
importance of geographically explicit interaction data.

@@

Here we demonstrated how we can detect uncertainty in species
distribution data using ecological interactions. Knowing where
questionable occurrence data are can be crucial in ecological modelling
(Hortal 2008; Ladle and Hortal 2013), and accounting for these errors
can improve model outputs by diminishing the error propagation (Draper
1995). For instance, we believe this is a way to account for ecological
interactions in habitat suitability models without making the models
more complex, but by making sure (not assuming) that the input data -
the species occurrence - actually accounts for ecological interactions.
It is important to notice, however, that the quality and usefulness of
this method are highly correlated with the amount and quality of data
available about species' occurrences \textbf{and} interactions. With
this paper, we hope to add to the collective effort to decode the
encrypted message that is the occurrence of a species in space and time.
A promising avenue that adds to our method is the prediction of networks
and interactions in large scales (Strydom et al. 2021; Windsor et al.
2022), for they can add valuable information about ecological
interactions where they are missing. Additionally, in order to achieve a
robust modelling framework towards actual species distribution models we
should invest in efforts to collect and combine open data on species
occurrence and interactions (Windsor et al. 2022), especially because we
may be losing ecological interactions at least as fast as we are losing
species (\textbf{Valiente-Banuet2015SpeLos?}).

@@

@@

>>>>

Finding a species in a certain location is like finding an encrypted
message that traveled through time. It carries the species' evolutionary
history, migration patterns, as well as any direct and indirect effects
generated by other species (some of which we may not even know exist).
Ecologists have been trying to decode this message with progressively
more powerful tools, from their field notes to highly complex
computational algorithms. However, to succeed in this challenge it is
important to have the right clues in hand. There are many ways we can be
misled by data - or the lack of it: taxonomic errors (e.g., due to
updates in the taxonomy of a species), geographic inaccuracy (e.g.,
approximate coordinates or lack of documentation about their accuracy),
or sampling biases (e.g.~data clustered near roads or research centers)
(Ladle and Hortal 2013; Hortal et al. 2015; Poisot et al. 2021). One way
to identify - and potentially fix - these errors is to combine many
different pieces of information about the occurrence of a species, so
agreements and mismatches can emerge. Although previous studies have
combined different types of occurrence data to measure the accuracy of
datasets (Hurlbert and Jetz 2007; Hurlbert and White 2005; Ficetola et
al. 2014), none have used different types of information so far (i.e.,
ecological characteristics other than geographical distribution). Here
we suggest jointly analysing species occurrence (range maps and point
occurrences) and ecological interactions to identify mismatches between
datasets.

@@

It has been demonstrated that the agreement between range maps and point
data varies geographically (Hurlbert and Jetz 2007; Hurlbert and White
2005; Ficetola et al. 2014). Adding ecological interaction data to this
comparison might help to elucidate where these (dis)agreements are more
likely to be true and which dataset better represent the actual
distribution of species. In this context, we elaborate a method that
allows us to refine distribution data (more precisely range maps) based
on interaction data, considering the basic assumption that predators can
only be present in regions where they are connected to at least one
herbivore - and thus indirectly connected to primary producers. We used
a Serengeti food web dataset (Baskerville et al. 2011) (which comprises
carnivores, herbivores, and plants from Tanzania) to demonstrate how a
mismatch between occurrence and interaction data can highlight
significant uncertainty areas in IUCN range maps. Finally, we add the
GBIF occurrence points for the Serengeti species to the investigation,
discuss the mechanisms that can lead to the lack of agreement between
data, and build from that a vision for the next steps, reinforcing the
importance of geographically explicit interaction data.

@@

Here we demonstrated how we can detect uncertainty in species
distribution data using ecological interactions. Knowing where
questionable occurrence data are can be crucial in ecological modelling
(Hortal 2008; Ladle and Hortal 2013), and accounting for these errors
can improve model outputs by diminishing the error propagation (Draper
1995). For instance, we believe this is a way to account for ecological
interactions in habitat suitability models without making the models
more complex, but by making sure (not assuming) that the input data -
the species occurrence - actually accounts for ecological interactions.
It is important to notice, however, that the quality and usefulness of
this method are highly correlated with the amount and quality of data
available about species' occurrences \textbf{and} interactions. With
this paper, we hope to add to the collective effort to decode the
encrypted message that is the occurrence of a species in space and time.
A promising avenue that adds to our method is the prediction of networks
and interactions in large scales (Strydom et al. 2021; Windsor et al.
2022), for they can add valuable information about ecological
interactions where they are missing. Additionally, in order to achieve a
robust modelling framework towards actual species distribution models we
should invest in efforts to collect and combine open data on species
occurrence and interactions (Windsor et al. 2022), especially because we
may be losing ecological interactions at least as fast as we are losing
species (Valiente-Banuet et al. 2015).

@@

\leavevmode\hypertarget{ref-Ficetola2014EvaRoba}{}%
Ficetola, Gentile Francesco, Carlo Rondinini, Anna Bonardi, Vineet
Katariya, Emilio Padoa-Schioppa, and Ariadne Angulo. 2014. {``An
Evaluation of the Robustness of Global Amphibian Range Maps.''}
\emph{Journal of Biogeography} 41 (2): 211--21.
\url{https://doi.org/10.1111/jbi.12206}.

@@

\leavevmode\hypertarget{ref-Valiente-Banuet2015SpeLosa}{}%
Valiente-Banuet, Alfonso, Marcelo A. Aizen, Julio M. Alcántara, Juan
Arroyo, Andrea Cocucci, Mauro Galetti, María B. García, et al. 2015.
{``Beyond Species Loss: The Extinction of Ecological Interactions in a
Changing World.''} Edited by Marc Johnson. \emph{Functional Ecology} 29
(3): 299--307. \url{https://doi.org/10.1111/1365-2435.12356}.
